Food-web complexity flattens the fitness landscape of an insect herbivore

### 2018-07-02 elsevier_test/elsevier_test.Rmd
updating plots and analyses in ms

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette

@@

## Custom functions to avoid repetitive code ----
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

gam.plot <- function(gam.model, ...) {
  plot(gam.model, se = TRUE, seWithMean = T, shift = mean(predict(gam.model)), trans = function(x) {exp(x)/(1+exp(x))}, ...)
}

# experimenting with boot.case gave bootstrapped distributions that weren't centered near the model mean estimate.
# also, using 'refit.smooth = T' also generated some very skewed distributions
# the recipe below provided the most sensible estimates for my data
gradient.calc <- function(mod, phenotype, covariates = NULL) {
  gam.gradients(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32, 
                refit.smooth = F,
                n.boot = 1000,
                se.method = 'boot.para', #'posterior', #'boot.para', #'permute', 'boot.case'#  
                standardized = T)
}

insect_FL <- function(mod, phenotype, covariates = NULL) {
  fitness.landscape(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32,
                    PI.method = 'boot.para',
                    refit.smooth = F,
                    PI.interval = c(0.025,0.975),
                    plt.density = 100)
}

get_insect_FL.df <- function(FL){
  data.frame(mean_fitness = FL$Wbar, FL$points, t(FL$WbarPI)) %>%
    mutate(lower_2.5 = X2.5., upper_97.5 = X97.5.)
}

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

inverse_logit <- function(x) exp(x)/(1+exp(x))

#Beta_avg_grad <- function(logistic_model, mean_fitness) {
#  brackets <- predict(logistic_model, type = "response") * (1 - predict(logistic_model, type="response"))
#  gradient <- mean(brackets) * fixef(logistic_model) / mean_fitness 
#  return(gradient)
#}

## Calculate directional selection gradients
# following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

## Bootstrap fitness estimates
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

@@

  # Organize data
  return(list(absolute_fitness = absolute_fitness_df, relative_fitness = relative_fitness_df))
}
```

@@

Biological diversity -- from genes, to phenotypes, to species -- has and continues to be shaped by the interplay between ecological and evolutionary processes. <!-- need to improve this instead of just saying its fascinated a small group of people --> 
Much of this biological diversity has been molded by natural selection arising from species interactions, such as resource competition (@Schluter2000), mutualisms (@Jordano1987), and predation (@Abrams2000). 
While there is clear evidence that pairwise interactions can drive evolution, we also know that most species interact with multiple species in an ecological community. <!-- are embedded within species-rich communities.--> <!-- complex networks of interacting species. <!-- need to make the link to multiple traits and the adaptive landscape here --> <!-- Somthing about the this work remaining eminently theoretical --> 
Understanding how evolutionary dynamics unfold in a community context<!-- when moving beyond pairwise interactions--><!-- in more species-rich communities--> is challenging and eminently theoretical (@McPeek2017, Nuismer paper, Guimeras paper, also McGill paper about biodiversity effects on evolution). 
<!-- Do I need to give more credit here to work by @TerHorst2018 and others?). I think it would be useful to show that there is some field and lab experiments demonstrating that community context can alter evolutionary trajectories -->
Given the rapid loss of species diversity we are experiencing throughout the world (cite), we are in urgent need of work that makes and tests predictions for how the loss of species will affect the evolutionary process in natural communities.

@@

Predicting the evolutionary consequences of biodiversity loss first requires an understanding of the concommitant change in the network structure of species interactions.<!-- the relationship between the network structure of a community and the fitness landscape of its constituent species.--> 
Knowing the interaction network is crucial because the loss of biodiversity, in and of itself, will not alter evolution -- it is the associated loss of ecological interactions that will affect evolutionary change (@Janzen1974). 
For a network of directly connected species, we would expect that the loss of species to result in a loss of network complexity.<!-- Need to formalize this logic more, because the loss of a specialist species in a highly generalist food web, would actually increase network complexity... Maybe I need to focus on the node/species level property --> 
Network complexity is a property that describes the diversity of interactions in an ecological community (@Banasek-Richter2009). 
Thus, all else equal, species loss will decrease the diversity of interactions, resulting in a more simple network. <!-- Explicitly linking network complexity to the fitness landscape, however, is ultimately needed to predict the evolutionary consequences of biodiversity loss. -->

@@

To characterize the shape of the fitness landscape, we quantified selection gradients acting on each trait in simple vs. complex food webs. We did this by fitting separate statistical models to data from each food-web treatment. We used generalized linear mixed models (GLMMs, @Bolker2009) with larval survival (0 or 1) as our response variable and measure of fitness. We specified linear and quadratic terms for each gall trait as well as linear interaction terms between each gall trait, as fixed effects in the statistical models. To account for the correlated structure of clutch size (gall level) and female preference (plant level) as well as any other independent effects of willow genotype on larval survival, we specified gall ID nested within plant ID nested within plant genotype as random intercepts in our statistical models. Since we were interested in characterizing the fitness landscape -- the relationship between mean trait values and population mean fitness -- we assumed the mean value of our random effects (i.e. setting them to zero) to estimate selection gradients. We then used the method of @Janzen1998 to calculate directional ($\beta_{z_i}$), nonlinear ($\gamma_{z_i}$), and correlational ($\gamma_{z_i,z_j}$) selection gradients and used parametric bootstrapping (1000 replicates) to calculate their 95% confidence intervals (@Bolker2009). To test whether selection gradients differed between treatments, we used our bootstrapped estimates to calculate the probability that selection gradients in the simple food web were larger/smaller than in the complex food web (i.e. the p-value). All analyses and visualizations were conducted in R.

```{r Selection Gradients - Complex Food Web, cache=TRUE, eval=FALSE}
# note that I've switched to using the lme4 package to analyzing the GLMM

# GLMM ----
glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                         I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2) +
                         sc.gall_size:sc.clutch_size + sc.gall_size:sc.female_preference + sc.clutch_size:sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = control_df,
                         family = binomial(link = logit), optimizer = "bobyqa")
# summary(glmer_control)

```{r Selection Gradients - Simple Food Web, cache=TRUE, eval=FALSE}

## GLMM ----
glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                         I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2) +
                         sc.gall_size:sc.clutch_size + 
                          sc.gall_size:sc.female_preference + 
                           sc.clutch_size:sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df,
                         family = binomial(link = logit), optimizer = "bobyqa")
summary(glmer_treatment)

## Mean fitness ----
Wbar_treatment <- inverse_logit(fixef(update(glmer_treatment, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

## Confidence intervals for selection gradients ----
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=100, parallel="multicore", ncpus=32)

## Summary ----
summary_Beta_glmm_treatment <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Simple",
  Model = "GLMM",
  Beta = round(Beta_avg_grad(glmer_treatment),2),
  lower_2.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
  upper_97.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2))

summary_Beta_glmm_treatment
```

@@

>>>>

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette
library(lme4) # for generalized linear mixed models

@@

## Custom functions to avoid repetitive code ----

## Test for overdispersion
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

## Transform logit to probability
inverse_logit <- function(x) exp(x)/(1+exp(x))

## Calculate selection gradients following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

## Bootstrap fitness estimates
bootstrap_fitness <- function(logistic_model, fixed_effects, newdata, bootstraps, intervals=c(0.025,0.975)){
  model_matrix <- model.matrix(formula(fixed_effects), newdata)

@@

  # Organize data
  return(list(absolute_fitness = absolute_fitness_df, relative_fitness = relative_fitness_df))
}

## Number of bootstrap replicates ----
n_boots_analysis <- 50
n_boots_plots <- 50
```

@@

Biological diversity -- from genes, to phenotypes, to species -- has and continues to be shaped by the interplay between ecological and evolutionary processes. <!-- need to improve this instead of just saying its fascinated a small group of people --> 
Much of this biological diversity has been molded by natural selection arising from species interactions, such as resource competition (@Schluter2000), mutualisms (@Jordano1987), and predation (@Abrams2000). 
While there is clear evidence that pairwise interactions can drive evolution, we also know that most species interact with multiple species in an ecological community. <!-- are embedded within species-rich communities.--> <!-- complex networks of interacting species. <!-- need to make the link to multiple traits and the adaptive landscape here --> <!-- Somthing about the this work remaining eminently theoretical --> 
Understanding how evolutionary dynamics unfold in a community context<!-- when moving beyond pairwise interactions--><!-- in more species-rich communities--> is challenging and eminently theoretical (@McPeek2017, @De_Mazancourt2008, @Guimaraes2017, @Nuismer2013). 
<!-- Do I need to give more credit here to work by @TerHorst2018 and others?). I think it would be useful to show that there is some field and lab experiments demonstrating that community context can alter evolutionary trajectories -->
Given the rapid loss of species diversity we are experiencing throughout the world (cite), we are in urgent need of work that makes and tests predictions for how the loss of species will affect the evolutionary process in natural communities.

@@

Predicting the evolutionary consequences of species loss first requires an understanding of the concommitant change in the species-interaction networks.<!-- the relationship between the network structure of a community and the fitness landscape of its constituent species.--> 
Knowing the interaction network is crucial because the loss of biodiversity, in and of itself, will not alter evolution -- it is the associated loss of ecological interactions that will affect evolutionary change (@Janzen1974). 
For a network of directly connected species, we would expect that the loss of species to result in a loss of network complexity.<!-- Need to formalize this logic more, because the loss of a specialist species in a highly generalist food web, would actually increase network complexity... Maybe I need to focus on the node/species level property --> 
Network complexity is a property that describes the diversity of interactions in an ecological community (@Banasek-Richter2009). 
Thus, all else equal, species loss will decrease the diversity of interactions, resulting in a more simple network. <!-- Explicitly linking network complexity to the fitness landscape, however, is ultimately needed to predict the evolutionary consequences of biodiversity loss. -->

@@

To characterize the shape of the fitness landscape, we quantified selection gradients acting on each trait in simple vs. complex food webs. We did this by fitting separate statistical models to data from each food-web treatment. We used generalized linear mixed models (GLMMs, @Bolker2009) with larval survival (0 or 1) as our response variable and measure of fitness. We specified linear and quadratic terms for each gall trait as well as linear interaction terms between each gall trait as fixed effects in the statistical models. To account for the correlated structure of clutch size (gall level) and female preference (plant level) as well as any other independent effects of willow genotype on larval survival, we specified gall ID nested within plant ID nested within plant genotype as random intercepts in our statistical models. Since we were interested in characterizing the fitness landscape -- the relationship between mean trait values and population mean fitness -- we assumed the mean value of our random effects (i.e. setting them to zero) to estimate selection gradients. We then used the method of @Janzen1998 to calculate directional ($\beta_{z_i}$), quadratic ($\gamma_{z_i}$), and correlational ($\gamma_{z_i,z_j}$) selection gradients and used parametric bootstrapping (1000 replicates) to calculate their 95% confidence intervals (@Bolker2009). To test whether selection gradients differed between treatments, we used our bootstrapped estimates to calculate the probability that selection gradients in the simple food web were larger/smaller than in the complex food web (i.e. the p-value). All analyses and visualizations were conducted in R (@R2018).

```{r Selection Gradients - Complex Food Web, cache=TRUE, include=FALSE}
# note that I've switched to using the lme4 package to analyzing the GLMM

# Directional selection ----
beta_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = control_df,
                         family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))

summary(beta_control)
overdisp_fun(beta_control)

# Quadratic selection ----
quad_control <- update(beta_control, .~. + I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2))

summary(quad_control)
overdisp_fun(quad_control)

# Correlational selection ----
corr_control <- update(quad_control, .~. + sc.gall_size:sc.clutch_size + sc.gall_size:sc.female_preference + sc.clutch_size:sc.female_preference)

summary(corr_control)
overdisp_fun(corr_control)

# Mean fitness ----
Wbar_control <- inverse_logit(fixef(update(beta_control, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Confidence intervals ----
boot_beta_control <- bootMer(beta_control, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
boot_quad_control <- bootMer(quad_control, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
boot_corr_control <- bootMer(corr_control, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
```

```{r Selection Gradients - Simple Food Web, cache=TRUE, include=FALSE}

# Directional selection ----
beta_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df,
                         family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))

summary(beta_treatment)
overdisp_fun(beta_treatment)

# Quadratic selection ----
quad_treatment <- update(beta_treatment, .~. + I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2))

summary(quad_treatment)
overdisp_fun(quad_treatment)

# Correlational selection ----
corr_treatment <- update(quad_treatment, .~. + sc.gall_size:sc.clutch_size + sc.gall_size:sc.female_preference + sc.clutch_size:sc.female_preference)

summary(corr_treatment)
overdisp_fun(corr_treatment)

# Mean fitness ----
Wbar_treatment <- inverse_logit(fixef(update(beta_treatment, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Confidence intervals ----
boot_beta_treatment <- bootMer(beta_treatment, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
boot_quad_treatment <- bootMer(quad_treatment, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
boot_corr_treatment <- bootMer(corr_treatment, FUN = Beta_avg_grad, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
```

```{r Table of Selection Gradients, echo=FALSE, message=FALSE, warning=FALSE}

## Summarize Control Selection Gradients ----
summary_beta_control <- data.frame(
  Phenotype = c("Gall diameter", "Clutch size", "Female preference"),
  Food_Web = "Complex",
  Model = "GLMM",
  Type = "Beta",
  Estimate = round(Beta_avg_grad(beta_control),3),
  lower_2.5 = round(apply(boot_beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3),
  upper_97.5 = round(apply(boot_beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3))

# doubled quadratic estimates to make comparable to directional selection gradients (Stinchcombe et al. 2008, Evolution)
summary_quad_control <- data.frame(
  Phenotype = c("Gall diameter", "Clutch size", "Female preference"),
  Food_Web = "Complex",
  Model = "GLMM",
  Type = "Quadratic",
  Estimate = 2*round(Beta_avg_grad(quad_control),3)[-c(1:3)],
  lower_2.5 = 2*round(apply(boot_quad_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3)[-c(1:3)],
  upper_97.5 = 2*round(apply(boot_quad_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3)[-c(1:3)])

summary_corr_control <- data.frame(
  Phenotype = c("Diam,Clutch", "Diam,Pref", "Clutch,Pref"),
  Food_Web = "Complex",
  Model = "GLMM",
  Type = "Correlational",
  Estimate = round(Beta_avg_grad(corr_control),3)[-c(1:6)],
  lower_2.5 = round(apply(boot_corr_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3)[-c(1:6)],
  upper_97.5 = round(apply(boot_corr_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3)[-c(1:6)])

## Summarize Treatment Selection Gradients ----
summary_beta_treatment <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Simple",
  Model = "GLMM",
  Type = "Beta",
  Estimate = round(Beta_avg_grad(beta_treatment),3),
  lower_2.5 = round(apply(boot_beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3),
  upper_97.5 = round(apply(boot_beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3))

# doubled quadratic estimates to make comparable to directional selection gradients (Stinchcombe et al. 2008, Evolution)
summary_quad_treatment <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Simple",
  Model = "GLMM",
  Type = "Quadratic",
  Estimate = 2*round(Beta_avg_grad(quad_treatment),3)[-c(1:3)],
  lower_2.5 = 2*round(apply(boot_quad_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3)[-c(1:3)],
  upper_97.5 = 2*round(apply(boot_quad_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3)[-c(1:3)])

summary_corr_treatment <- data.frame(
  Phenotype = c("Diam,Clutch", "Diam,Pref", "Clutch,Pref"),
  Food_Web = "Simple",
  Model = "GLMM",
  Type = "Correlational",
  Estimate = round(Beta_avg_grad(corr_treatment),3)[-c(1:6)],
  lower_2.5 = round(apply(boot_corr_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3)[-c(1:6)],
  upper_97.5 = round(apply(boot_corr_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3)[-c(1:6)])

bind_rows(summary_beta_control, summary_beta_treatment)
bind_rows(summary_quad_control, summary_quad_treatment)
bind_rows(summary_corr_control, summary_corr_treatment)
```

@@

BREAK UP INTO INDIVIDUAL PIECES: E.G. GALL SIZE. THIS SHOULD BE SMALL ENOUGH TO ALLOW CACHING

```{r Get Univariate Fitness Landscapes, echo=FALSE, message=FALSE, warning=FALSE}
## Gall size ----

# Control
RF_control_gall_size <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000), 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

# Treatment
RF_treatment_gall_size <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000),
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

## Clutch size ----

# Control
RF_control_clutch_size <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(-1,1,length.out=1000),
     sc.female_preference = 0),
  bootstraps=100)

# Treatment 
RF_treatment_clutch_size <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(-1,1,length.out=1000),
     sc.female_preference = 0),
  bootstraps=100)

## Female preference ----

# Control 
RF_control_female_preference <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0, 
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=100)

# Treatment
RF_treatment_female_preference <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0,
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=100)
```

```{r Plot Univariate Landscapes, echo=FALSE, message=FALSE, warning=FALSE}
## Relative fitness plots ----

# Get range in relative fitness for accurate scaling across all plots
RF_gradient_range <- bind_rows(RF_treatment_gall_size$relative_fitness, RF_control_gall_size$relative_fitness,
          RF_treatment_clutch_size$relative_fitness, RF_control_clutch_size$relative_fitness,
          RF_treatment_female_preference$relative_fitness, RF_control_female_preference$relative_fitness) %>%
  select(-sc.clutch_size, -sc.female_preference) %>%
  gather(ID, relative_fitness, -sc.gall_size) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  summarise(min = min(relative_fitness), max = max(relative_fitness))
RF_gradient_scale <- c(0,0.5,1,1.5,2,2.5) # adjusted based on range

# Color palette: order -> orange, light blue, green, yellow, dark blue, red, pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#999999")
complex_color <- cbPalette[6]
simple_color <- cbPalette[5]
treatment_colors <- c(complex_color, simple_color)

# Gall size    
RF_uni_size <- bind_rows(mutate(RF_treatment_gall_size$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_gall_size$relative_fitness, Food_Web = "Complex")) %>%
  select(-sc.clutch_size, -sc.female_preference) %>%
  gather(ID, relative_fitness, -sc.gall_size, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.gall_size, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_alpha_manual(values=c(1,0.1), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Gall diameter (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_y_continuous(breaks = RF_gradient_scale, limits = c(min(RF_gradient_scale), max(RF_gradient_scale)))

# Clutch size
RF_uni_clutch <- bind_rows(mutate(RF_treatment_clutch_size$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_clutch_size$relative_fitness, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.female_preference) %>%
  gather(ID, relative_fitness, -sc.clutch_size, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.clutch_size, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_alpha_manual(values=c(1,0.1), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Clutch size (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_y_continuous(breaks = RF_gradient_scale, limits = c(min(RF_gradient_scale), max(RF_gradient_scale)))

# Female preference 
RF_uni_pref <- bind_rows(mutate(RF_treatment_female_preference$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_female_preference$relative_fitness, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.clutch_size) %>%
  gather(ID, relative_fitness, -sc.female_preference, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.female_preference, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_alpha_manual(values=c(1,0.1), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Female preference (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_y_continuous(breaks = RF_gradient_scale, limits = c(min(RF_gradient_scale), max(RF_gradient_scale)))

# Get legend
RF_uni_legend <- get_legend(RF_uni_size)

# Make plots
RF_uni_plots <- plot_grid(RF_uni_size + theme(legend.position = "none"), 
                       RF_uni_clutch + theme(legend.position = "none", 
                                          axis.title.y = element_blank(), axis.text.y = element_blank()),
                       RF_uni_pref + theme(legend.position = "none", 
                                        axis.title.y = element_blank(), axis.text.y = element_blank()),
                       labels = "AUTO", ncol = 4, align='hv')

RF_gradients <- ggdraw(RF_uni_plots) + draw_grob(RF_uni_legend, x = 0.8, y=0)
RF_gradients
#save_plot("RF_gradients.png", RF_gradients, base_height=6, base_width = 8)
#save_plot("RF_gradients.pdf", RF_gradients, base_height=6, base_width = 8)
```

```{r Get Multivariate Landscapes, echo=FALSE, message=FALSE, warning=FALSE}

## Clutch size x Female preference ----

# Control
RF_control_clutch.x.preference <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(-1,1,length.out=1000), 
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=NULL)

# Treatment
RF_treatment_clutch.x.preference <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(-1,1,length.out=1000),
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=NULL)

## Gall size x Clutch size ----

# Control
RF_control_size.x.clutch <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000),
     sc.clutch_size = seq(-1,1,length.out=1000), 
     sc.female_preference = 0), 
  bootstraps=NULL)

# Treatment
RF_treatment_size.x.clutch <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000),
     sc.clutch_size = seq(-1,1,length.out=1000),
     sc.female_preference = 0), 
  bootstraps=NULL)

## Gall size x Female preference ----

# Control
RF_control_size.x.preference <- bootstrap_fitness(
  logistic_model = beta_control, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000),
     sc.clutch_size = 0, 
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=NULL)

# Treatment
RF_treatment_size.x.preference <- bootstrap_fitness(
  logistic_model = beta_treatment, 
  fixed_effects = "~ sc.gall_size + sc.clutch_size + sc.female_preference",
  newdata = expand.grid(
     sc.gall_size = seq(-1,1,length.out=1000),
     sc.clutch_size = 0,
     sc.female_preference = seq(-1,1,length.out=1000)), 
  bootstraps=NULL)
```

```{r Plot Multivariate Fitness Landscape, echo=FALSE, message=FALSE, warning=FALSE}

## Relative fitness plots ----

# Get range in relative fitness for accurate scaling across all plots
RF_range <- bind_rows(RF_treatment_clutch.x.preference$relative_fitness, RF_control_clutch.x.preference$relative_fitness,
          RF_treatment_size.x.clutch$relative_fitness, RF_control_size.x.clutch$relative_fitness,
          RF_treatment_size.x.preference$relative_fitness, RF_control_size.x.preference$relative_fitness) %>%
  summarise(min = min(average), max = max(average))

# Clutch x Preference
RF_plot_clutch.x.preference <- bind_rows(mutate(RF_treatment_clutch.x.preference$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_clutch.x.preference$relative_fitness, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.female_preference, y=sc.clutch_size, fill=average)) +
  geom_raster() +
  facet_wrap(~Food_Web) +
  xlab("Female preference (SDs)") +
  ylab("Clutch size (SDs)") +
  #scale_fill_gradient2(midpoint=1, limits = c(RF_range$min, RF_range$max))
  scale_fill_viridis(limits = c(RF_range$min, RF_range$max), name = "Relative fitness")

# Size x Clutch
RF_plot_size.x.clutch <- bind_rows(mutate(RF_treatment_size.x.clutch$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_size.x.clutch$relative_fitness, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.gall_size, y=sc.clutch_size, fill=average)) +
  geom_raster() +
  facet_wrap(~Food_Web) +
  xlab("Gall diameter (SDs)") +
  ylab("Clutch size (SDs)") +
  #scale_fill_gradient2(midpoint=1, limits = c(RF_range$min, RF_range$max))
  scale_fill_viridis(limits = c(RF_range$min, RF_range$max), name = "Relative fitness")

# Size x Preference
RF_plot_size.x.preference <- bind_rows(mutate(RF_treatment_size.x.preference$relative_fitness, Food_Web = "Simple"), 
          mutate(RF_control_size.x.preference$relative_fitness, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.gall_size, y=sc.female_preference, fill=average)) +
  geom_raster() +
  facet_wrap(~Food_Web) +
  xlab("Gall diameter (SDs)") +
  ylab("Female preference (SDs)") +
  #scale_fill_gradient2(midpoint=1, limits = c(RF_range$min, RF_range$max))
  scale_fill_viridis(limits = c(RF_range$min, RF_range$max), name = "Relative fitness")

# Get legend
RF_landscape_legend <- get_legend(RF_plot_size.x.preference)
